---
output:
  github_document:
    html_preview: TRUE
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "## ",
  fig.path = "README-"
)
```

# ShortForm

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version-last-release/ShortForm)](http://cran.r-project.org/package=ShortForm) [![R-CMD-check](https://github.com/AnthonyRaborn/ShortForm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/AnthonyRaborn/ShortForm/actions/workflows/R-CMD-check.yaml) [![CRAN Downloads Per Month](https://cranlogs.r-pkg.org/badges/ShortForm)](https://cran.r-project.org/package=ShortForm) [![CRAN Downloads Total](https://cranlogs.r-pkg.org/badges/grand-total/ShortForm?color=orange)](https://cran.r-project.org/package=ShortForm)

## Overview

**ShortForm** is an R package for constructing short-form assessments from larger item banks using reproducible, optimization-based workflows.

It provides implementations of metaheuristic search algorithms to automate item selection while preserving prespecified psychometric properties. The package is designed for researchers and practitioners who need scalable, transparent, and reproducible methods for reducing assessment length.

------------------------------------------------------------------------

## Key Features

-   Automated short-form construction from full item banks
-   Multiple optimization algorithms:
    -   Ant Colony Optimization (ACO)
    -   Tabu Search (TS)
    -   Simulated Annealing (SA)
-   Integration with `lavaan` and `Mplus` workflows
-   Support for custom objective functions and model specifications
-   Reproducible, scriptable workflows for assessment design

------------------------------------------------------------------------

## Use Cases

ShortForm is useful when you need to:

-   Reduce assessment length while maintaining measurement validity
-   Automate item selection workflows for large-scale assessments
-   Build reproducilbe pipelines for psychometric modeling and evaluation

------------------------------------------------------------------------

## Engineering Highlights

This package demonstrates:

-   Development and maintenance of an R package distributed via CRAN
-   Implementation of optimization algorithms for applied statistical workflows
-   Automated build and testing pipelines using CI/CD
-   Integration with other modeling tools
-   Reproducible examples using structured simulated data

------------------------------------------------------------------------

This document was created on `r Sys.Date()`.

## Installation

```{r gh-installation, eval = FALSE}
install.packages("ShortForm") # the CRAN-approved version
require("devtools")
devtools::install_github("AnthonyRaborn/ShortForm", branch = "devel") # the developmental version
```

## Quick Examples

```{r quick-example-ACO}
set.seed(58310)
library(ShortForm)

# Example workflow:
# 1. Define item pool and model
# 2. Run optimization algorithm
# 3. Evaluate resulting short form

result_ACO <- antcolony.lavaan(
    data = simulated_test_data,
    antModel = exampleAntModel,
    list.items = list(paste0("Item", 1:56)), 
    full = 56, i.per.f = 15,
    factors = "Ability", 
    steps = 100,
    fit.indices = c("cfi", "rmsea"),
    fit.statistics.test = "(cfi > 0.98)&(rmsea < 0.03)",
    summaryfile = NULL,
    feedbackfile = NULL,
    max.run = 200
)

summary(result_ACO)
plot(result_ACO)
```

```{r quick-example-SA}
set.seed(58310)
result_SA <- simulatedAnnealing(
  initialModel = paste("Ability =~", paste0("Item", 1:56, collapse = " + ")),
  originalData = simulated_test_data,
  maxSteps = 20, 
  maxItems = 40, 
  items = paste0("Item", 1:56),
  setChains = 4
)

summary(result_SA)
plot(result_SA)
```

```{r quick-example-TS}
set.seed(58310)
shortAntModel <- "
Ability =~ Item1 + Item2 + Item3 + Item4 + Item5 + Item6 + Item7 + Item8
Ability ~ Outcome
"

result_TS <- tabuShortForm(
  initialModel = shortAntModel,
  originalData = simulated_test_data, numItems = 7,
  niter = 1, tabu.size = 3, parallel = FALSE
)
summary(result_TS)
```

## Workflow

```{mermaid}
flowchart LR
    A@{ shape: lean-r, label: "Full Item Bank" } --> B[Specify Model]
    B --> C[Run Optimization Algorithm]
    C --> D[Evaluate Candidate Forms]
    D --> |One or More Forms Fit| E@{ shape: diamond, label: "Select Short Form" }
    D --> |No Adequate Forms Found| C
    E --> F@{ shape: lean-l, label: "Reporting & Visualization" }]
    
```

## Outputs

ShortForm produces diagnostic outputs such as:

-   Objective function convergence plots
-   Algorithm-specific diagnostics
-   Summary statistics for selected item sets

------------------------------------------------------------------------

## Algorithms

### Ant Colony Optimization (ACO)

Implementation adapted from [Leite, Huang, & Marcoulides (2008)](https://doi.org/10.1080/00273170802285743).

### Tabu Search (TS)

Based on [Marcoulides & Falk (2018)](https://doi.org/10.1080/10705511.2017.1409074), extended for model search and short-form construction.

### Simulated Annealing (SA)

Provides one of the first implementations of simulated annealing for psychometric model selection workflows. Implemented following [Kirkpatrick et al. (1983)](https://doi.org/10.1126/science.220.4598.671).

------------------------------------------------------------------------

## Performance Notes

Computational time depends on: 

- Number and types of items 
- Dimensionality/model complexity 
- Algorithm convergence settings

Parallelization is supported for certain workflows.

------------------------------------------------------------------------

## Reproducibility

ShortForm is designed for reproducible analytical workflows:

- Fully scriptable 
- Compatible with R Markdown and Quarto 
- Supports integration into larger modeling pipelines

------------------------------------------------------------------------

## License

LGPL (\>= 2.0, \< 3) \| Mozilla Public License

------------------------------------------------------------------------

## Citation

If you use this package, please cite:

Raborn A, Leite W (2024). *ShortForm: Automatic Short Form Creation*. <doi:10.32614/CRAN.package>. ShortForm <https://doi.org/10.32614/CRAN.package.ShortForm>, R package version 0.5.6, <https://CRAN.R-project.org/package=ShortForm>.
