---
output:
  github_document:
    html_preview: false
---
```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ShortForm
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ShortForm)](http://cran.r-project.org/package=ShortForm)
[![Travis-CI Build Status](http://travis-ci.org/AnthonyRaborn/ShortForm.svg?branch=master)](http://travis-ci.org/AnthonyRaborn/ShortForm)

Automatic Short Form Creation for scales. Currently, only the Ant Colony Algorithm is implemented. The original R imnplementation for this algorithm is from Leite, Huang, & Marcoulides (2008) <doi:10.1080/00273170802285743>.

## Installation
```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("AnthonyRaborn/ShortForm") # the developmental version
install.packages("ShortForm") # the CRAN-approved version
```

## Usage

Here is a (slightly modified) example from the help documentation using lavaan.

```{r example, eval = FALSE}
# using simulated test data and the default values for lavaan.model.specs
# (with one exception), fit a 20-item short form
# first, read in the original or "full" model
data(exampleAntModel) # a character vector for a lavaan model

# then, create the list of the items by the factors
# in this case, all items load onto the general 'Ability' factor
list.items <- list(c('Item1','Item2','Item3','Item4','Item5',
'Item6','Item7','Item8','Item9','Item10',
'Item11','Item12','Item13','Item14','Item15',
'Item16','Item17','Item18','Item19','Item20',
'Item21','Item22','Item23','Item24','Item25',
'Item26','Item27','Item28','Item29','Item30',
'Item31','Item32','Item33','Item34','Item35',
'Item36','Item37','Item38','Item39','Item40',
'Item41','Item42','Item43','Item44','Item45',
'Item46','Item47','Item48','Item49','Item50',
'Item51','Item52','Item53','Item54','Item55','Item56'))

# load the data
data(simulated_test_data)
# finally, call the function with some minor changes to the default values.
# since the data is binary, let lavaan know by putting the items in the
# 'orderd' element of the lavaan.model.specs list.
abilityShortForm = antcolony.lavaan(data = simulated_test_data,
ants = 5, evaporation = 0.7, antModel = exampleAntModel,
list.items = list.items, full = 56, i.per.f = 20,
lavaan.model.specs = list(model.type = "cfa", auto.var = T, estimator = "default", ordered = unlist(list.items), int.ov.free = TRUE, int.lv.free = FALSE, auto.fix.first = TRUE, auto.fix.single = TRUE, auto.cov.lv.x = TRUE, auto.th = TRUE, auto.delta = TRUE, auto.cov.y = TRUE)
factors = 'Ability', steps = 3, fit.indices = c('cfi', 'rmsea'),
fit.statistics.test = "(cfi > 0.95)&(rmsea < 0.05)",
summaryfile = 'summary.txt',
feedbackfile = 'iteration.html',
max.run = 500)
abilityShortForm[[1]] # print the results of the final short form
```

A similar example can be found in the `antcolony.mplus` function, but requires you to have a valid Mplus installation on the computer.