---
output:
  github_document:
    html_preview: TRUE
---
```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "## ",
  fig.path = "README-"
)
```

# ShortForm
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ShortForm)](http://cran.r-project.org/package=ShortForm)
[![Travis-CI Build Status](http://travis-ci.org/AnthonyRaborn/ShortForm.svg?branch=master)](http://travis-ci.org/AnthonyRaborn/ShortForm)
[![CRAN Downloads Per Month](https://cranlogs.r-pkg.org/badges/ShortForm)](https://cran.r-project.org/package=ShortForm)
[![CRAN Downloads Total](https://cranlogs.r-pkg.org/badges/grand-total/ShortForm?color=orange)](https://cran.r-project.org/package=ShortForm)


Automatic Short Form Creation for scales. Currently, the Ant Colony Optimization  (ACO) Algorithm and the Tabu search are implemented. The original R implementation for the ACO algorithm is from Leite, Huang, & Marcoulides (2008) <doi:10.1080/00273170802285743>, while the Tabu search function was taken from Marcoulides & Falk (2018) <doi:10.1080/10705511.2017.1409074>.

## Installation
```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("AnthonyRaborn/ShortForm") # the developmental version
install.packages("ShortForm") # the CRAN-approved version
```

## Usage

Here are some (slightly modified) examples from the help documentation using
lavaan. Be warned, the algorithms may take some time to converge, particularly
with large forms, multiple dimensions, and different settings. The time for these
examples to converge on a low-end laptop is printed at the bottom.

### ACO Algorithm

```{r ACO example}
start.time.ACO <- Sys.time()
library(ShortForm)
# using simulated test data and the default values for lavaan.model.specs
# (with one exception), fit a 10-item short form
# first, read in the original or "full" model
data(exampleAntModel) # a character vector for a lavaan model

# then, create the list of the items by the factors
# in this case, all items load onto the general 'Ability' factor
list.items <- list(c('Item1','Item2','Item3','Item4','Item5',
'Item6','Item7','Item8','Item9','Item10',
'Item11','Item12','Item13','Item14','Item15',
'Item16','Item17','Item18','Item19','Item20',
'Item21','Item22','Item23','Item24','Item25',
'Item26','Item27','Item28','Item29','Item30'))

# load the data
data(simulated_test_data)
# finally, call the function with some minor changes to the default values.
# since the data is binary, let lavaan know by putting the items in the
# 'ordered' element of the lavaan.model.specs list.
set.seed(1)
abilityShortForm = antcolony.lavaan(data = simulated_test_data,
ants = 5, evaporation = 0.7, antModel = exampleAntModel,
list.items = list.items, full = 30, i.per.f = 10,
lavaan.model.specs = list(model.type = "cfa", auto.var = T, estimator = "default", 
                          ordered = unlist(list.items), int.ov.free = TRUE,
                          int.lv.free = FALSE, auto.fix.first = TRUE, 
                          auto.fix.single = TRUE, auto.cov.lv.x = TRUE, 
                          auto.th = TRUE, auto.delta = TRUE, 
                          auto.cov.y = TRUE),
factors = 'Ability', steps = 3, fit.indices = c('cfi', 'rmsea'),
fit.statistics.test = "(cfi > 0.90)&(rmsea < 0.10)",
summaryfile = NULL,
feedbackfile = NULL,
max.run = 50, verbose = FALSE)
abilityShortForm[[1]] # print the results of the final short form
```

A similar example can be found in the `antcolony.mplus` function, but requires you to have a valid Mplus installation on the computer. It took a total of `r round(as.difftime(Sys.time() - start.time.ACO, units = "mins"),2)` minutes to run this example.

### Tabu Search Algorithm

This example demonstrates how to use the Tabu search for model specification searches when the original model may be misspecified in some way.

```{r Tabu example}
start.time.Tabu <- Sys.time()
library(ShortForm)
# load simulation data and select columns used in this example
data(simulated_test_data) 
tabuData <- simulated_test_data[,c(1:10)]

# specify an improper model (improper because the data is actually unidimensional)
tabuModel <- "
Ability =~ Item1 + Item2 + Item3 + Item4
FakeAbility =~ Item5 + Item6 + Item7 + Item8
Ability ~ Outcome
FakeAbility ~ 0*Outcome
Ability ~ 0*FakeAbility"

# fit the initial misspecified model for Tabu
init.model <- lavaan::lavaan(model = tabuModel, data = tabuData, 
auto.var=TRUE, auto.fix.first=FALSE, std.lv=TRUE,auto.cov.lv.x=TRUE)

# use search.prep to prepare for the Tabu search
ptab <- search.prep(fitted.model = init.model, loadings=TRUE, fcov=TRUE, errors=FALSE)

# perform the Tabu search with 100 iterations and a Tabu list size of 5
set.seed(1) # reproduceable
Tabu_example <- suppressWarnings(tabu.sem(init.model = init.model, ptab = ptab, obj = AIC, niter = 25, tabu.size = 5)) # the suppressWarning wrapping hides the lavaan WARNING output from improper models

# check the final model
lavaan::summary(Tabu_example$best.mod)
```

It took a total of `r round(as.difftime(Sys.time() - start.time.Tabu, units = "mins"),2)` minutes to run this example.

### Simulated Annealing

This example demonstrates the use of simulated annealing for creating short forms.

```{r Simulated Annealing example}
start.time.SA <- Sys.time()
library(ShortForm)
# load simulation data and select columns used in this example
data(simulated_test_data) 
saData <- simulated_test_data[,c(1:10)]

# specify the full model
saModel <- "
Ability =~ Item1 + Item2 + Item3 + Item4 + Item5 + Item6 + Item7 + Item8 + Item9 + Item10
Ability ~ Outcome
"
lavaan.model.specs = list(model.type = "cfa",
  auto.var = TRUE, estimator = "default", ordered = paste0("Item", 1:10), int.ov.free = TRUE,
  int.lv.free = FALSE, std.lv = TRUE, auto.fix.first = FALSE, auto.fix.single =
  TRUE, auto.cov.lv.x = TRUE, auto.th = TRUE, auto.delta = TRUE, auto.cov.y =
  TRUE)

# perform the SA algorithm
set.seed(1)
SA_example <- simulatedAnnealing(initialModel = saModel, originalData = saData, maxSteps = 200, fitStatistic = 'cfi', maximize = FALSE, temperature = "logistic", items = paste0("Item", 1:10), lavaan.model.specs = lavaan.model.specs, maxChanges = 3, maxItems = 5, progressBar = F)
plot(SA_example$allFit, type = "l") # plot showing how the fit value changes at each step
```

It took a total of `r round(as.difftime(Sys.time() - start.time.SA, units = "mins"),2)` minutes to run the SA example, and a total of `r round(as.difftime(Sys.time() - start.time.ACO, units = "mins"),2)` minutes to run all three together.